---
title: "Lab: Reviewing Imported Data"
output: html_document
# Adapted from Carrie Wright and Ava Hoffman
---

Prompt: Read in the data sets and explore using various tools. Note, you may need to install the suggested tool packages first.


```{r, message = FALSE}
library(dplyr)
library(tidyverse)
library(naniar)
library(visdat)
```


1. Read in the the Charm City Circulator Dataset from `data/Charm_City_Circulator_Ridership.csv` using `read_csv`, call it `circ`
```{r}
Circ <- read_csv("C:/Users/14699/Desktop/UTSWMC/School_of_Health_Professions/Department_of_Health_Informatics/MSHI/My_Spring_2025_Courses/HI5305/Homeworks_Assignments/Dr_Andrew/cloud/RProject/data/raw_data/Charm_City_Circulator_Ridership.csv")
```

2. Print the number of rows of `circ`
```{r}
nrow(Circ)
```

3. Visualize the missing data patterns of `circ` with `vis_miss`:
```{r}
install.packages("visdat")
library(visdat)
vis_miss(Circ)
```

4. Determine the percent missing in each column using `miss_var_summary`

```{r}
install.packages("naniar")
library(naniar)
miss_var_summary(Circ)
```

5. Determine the number missing in each column using `gg_miss_var`:
```{r}
# Visualize the number of missing values in each column
gg_miss_var(Circ)
```

6. Write out `circ` to a file called "Circulator.csv" using `write_csv`:
```{r}
write_csv(Circ, "Circulator.csv")
```

7. Write out `circ` to a file called "Circulator.rds" using `write_rds`:

```{r}
write_rds(Circ, "Circulator.rds")
```