---
title: 'Lab: Data Wrangling Review 1'
output:
  pdf_document: default
editor_options:
  chunk_output_type: console
---

This lab will simply use the `mpg` dataset included with ggplot2.


```{r, message = FALSE}
library(tidyverse)
```

# Part 1

1. Check to see if you have the `mpg` dataset.

```{r}
data = mpg
head(mpg)
```


2. What data class is `mpg`?
```{r}
class(mpg)
```

3. How many observations (rows) and variables (columns) are in the `mpg` dataset?

```{r}
# Install and load necessary package 
install.packages("ggplot2")
library(ggplot2)

# Access the mpg dataset
mpg <- ggplot2::mpg
# It has 234 rows and 11 columns
```

4. Select the `manufacturer`, `model` and `year` columns from the `mpg` dataset.

```{r}
selected_data <- subset(mpg, select = c(manufacturer, model, year))
# OR
selected_data <- mpg[, c("manufacturer", "model", "year")]
View(selected_data)
```

5. Identify the subset of cars/rows where city fuel economy (`cty`) is greater than 20 and highway fuel economy (`hwy`) is greater than 30. Assign this output to an object called `eff`. How many cars/rows are present?

```{r}
eff <- subset(mpg, cty > 20 & hwy > 30) 
   # Or using logical indexing
eff <- mpg[mpg$cty > 20 & mpg$hwy > 30, ]
nrow(eff)
```

6. How many fuel efficient cars (in the `eff` dataset) were manufactured in the year 1999?

```{r}
eff_1999 <- subset(eff, year == 1999)
   # Or using logical indexing
eff_1999 <- eff[eff$year == 1999, ]
nrow(eff_1999)
```

7. Filter cars from the overall `mpg` dataset that do not have 4 cylinder engines. How many cars/rows are there?

```{r}
not_4cyl <- subset(mpg, cyl != 4)
   # Or using logical indexing
not_4cyl <- mpg[mpg$cyl != 4, ]
nrow(not_4cyl)
```

8. Filter cars to only those in the "suv" or "minivan" class. How many cars/rows are there?

```{r}
suv_minivan <- subset(mpg, class %in% c("suv", "minivan"))
   # Or using logical indexing
suv_minivan <- mpg[mpg$class %in% c("suv", "minivan"), ]
nrow(suv_minivan)
```

9. Filter cars with displacements (`displ`) greater than 4 and that are all 4 wheel drive (`drv`) (a value of 4 for `drv`). How many cars/rows are there? 

```{r}
filtered_cars <- subset(mpg, displ > 4 & drv == "4")
   # Or using logical indexing
filtered_cars <- mpg[mpg$displ > 4 & mpg$drv == "4", ]
nrow(filtered_cars)
```

