---
title: "Regex Lab"
author: "Jeff Leek"
date: "July 8, 2015"
output: html_document
---


1. Create a regular expression that finds any sentence with SISBID regardless of capitalization. To help you test it here is a character you can use with `str_detect`: 

```{r}
# x = c("SISBID","SiSBID","sisbid","sisBID")
regex("sisbid", ignore_case = TRUE)
# using str_detect():
library(stringr)

x <- c("SISBID", "SiSBID", "sisbid", "sisBID")

str_detect(x, regex("sisbid", ignore_case = TRUE))
# Output: TRUE TRUE TRUE TRUE; All entries match!
```

2. You can use `{n}` to match the previous pattern exactly n times. Try to create an expression that finds any statement with exactly three repeated 3's in it. To help you test it here is a character you can use with `str_detect`:

```{r}
# x = c(333,33,3,33333)
# a regular expression that detects exactly three 3's (333) — not more, not less.
"(^|[^3])3{3}([^3]|$)"
# This means: 3{3}: exactly three 3s; (^|[^3]): ensures the match starts at the beginning or after a non-3; ([^3]|$): ensures the match ends before a non-3 or at the end

x <- c("333", "33", "3", "33333")

str_detect(x, "(^|[^3])3{3}([^3]|$)")
# Output: TRUE FALSE FALSE FALSE; Only "333" is matched
```

3. You can use `{n,m}` to match a pattern between n and m times or `{n,}` to match a pattern at least n times. Try to create an expression that finds between 2 and 4 repeated f's. To help you test it here is a character you can use with `str_detect`:

```{r}
# x = c("jeff","jeffff","jef","je","j","jefffffff")
"f{2,4}" # This matches at least 2 and at most 4 consecutive fs
library(stringr)

x <- c("jeff", "jeffff", "jef", "je", "j", "jefffffff")

str_detect(x, "f{2,4}")
# Output: TRUE  TRUE FALSE FALSE FALSE  TRUE 
```

4. There are also character classes built in that you can use. For example "[:punct:]" finds all punctuation and "[:alpha:]" is an alphanumeric character. Try to create an expression that matches the pattern of a letter followed by punctuation at the end of the line. To help you test it here is a character you can use with `str_detect`:

```{r}
# x = c("yes.","no","why1",",ok")
"[[:alpha:]][[:punct:]]$"
# This means: [[:alpha:]]: a letter (A–Z or a–z); [[:punct:]]: a punctuation character; $: end of the line
x <- c("yes.", "no", "why1", ",ok")

str_detect(x, "[[:alpha:]][[:punct:]]$")
# Output: TRUE FALSE FALSE FALSE: Only "yes." matches — it ends with a letter + punctuation.
```

