--
title: "Lab: Data Importing with readr"
output: html_document
# Adapted from Carrie Wright and Ava Hoffman
---

```{r, include = TRUE}
library(readr)
library(readxl)
```

This lab uses different versions of the `iris` dataset. For each question, click the `iris` file in the Files pane to open it in an RStudio window (by selecting View File) . The files are in the `data/iris` folder. To open it, you can c

You can see more about this dataset by running `?iris` in the console.

1.  Read in the `iris_q1.csv` dataset into the `iris_q1` R object. What delimiter separates columns?

```{r}
iris_q1 <- read_csv("C:/Users/14699/Desktop/UTSWMC/School_of_Health_Professions/Department_of_Health_Informatics/MSHI/My_Spring_2025_Courses/HI5305/Homeworks_Assignments/Dr_Andrew/cloud/RProject/data/raw_data/iris/iris_q1.csv")
# The delimiter separating the columns is a comma (,).
```

2.  Read in the `iris_q2.tsv` dataset into the `iris_q2` R object. What delimiter separates columns?

```{r}
iris_q2 <- read_csv("C:/Users/14699/Desktop/UTSWMC/School_of_Health_Professions/Department_of_Health_Informatics/MSHI/My_Spring_2025_Courses/HI5305/Homeworks_Assignments/Dr_Andrew/cloud/RProject/data/raw_data/iris/iris_q2.tsv")
# The columns in the file are separated by a tab (\t) delimiter. 
```

3.  Read in the `iris_q3.txt` dataset into the `iris_q3` R object. What delimiter separates columns?

```{r}
iris_q3 <- read_csv("C:/Users/14699/Desktop/UTSWMC/School_of_Health_Professions/Department_of_Health_Informatics/MSHI/My_Spring_2025_Courses/HI5305/Homeworks_Assignments/Dr_Andrew/cloud/RProject/data/raw_data/iris/iris_q3.txt")
# It uses a space as the delimiter between columns.
```

4.  Read in the `iris_q4.csv` dataset into the `iris_q4` R object. What delimiter separates columns? [hint: file extension is ambiguous]

```{r}
iris_q4 <- read_csv("C:/Users/14699/Desktop/UTSWMC/School_of_Health_Professions/Department_of_Health_Informatics/MSHI/My_Spring_2025_Courses/HI5305/Homeworks_Assignments/Dr_Andrew/cloud/RProject/data/raw_data/iris/iris_q4.csv")
# The columns in the "iris_q4.csv" file are separated by a colon (:) delimiter. 
```

5.  Read in the `iris_q5.csv` dataset into the `iris_q5` R object. Some entries read "-999" but are supposed to be missing (the value `NA` in the data).

```{r}
iris_q5 <- read_csv("C:/Users/14699/Desktop/UTSWMC/School_of_Health_Professions/Department_of_Health_Informatics/MSHI/My_Spring_2025_Courses/HI5305/Homeworks_Assignments/Dr_Andrew/cloud/RProject/data/raw_data/iris/iris_q5.csv")
# A total NAs in the column of Sepal length
```

6.  Read in sheet 2 of the `iris_q6.xlsx` dataset into the `iris_q6` R object. How many rows are in the dataset?

```{r}
iris_q6 <- read_excel("C:/Users/14699/Desktop/UTSWMC/School_of_Health_Professions/Department_of_Health_Informatics/MSHI/My_Spring_2025_Courses/HI5305/Homeworks_Assignments/Dr_Andrew/cloud/RProject/data/raw_data/iris/iris_q6.xlsx")
# There are 150 rows                     
```
