---
title: 'Lab: Data Wrangling Review 1'
output:
  pdf_document: default
editor_options:
  chunk_output_type: console
# Adapted from Carrie Wright and Ava Hoffman
---

This lab will simply use the `mpg` dataset included with ggplot2.


```{r, message = FALSE}
library(tidyverse)
```

# Part 1

1. Check to see if you have the `mpg` dataset.

```{r}



```


2. What data class is `mpg`?
```{r}



```

3. How many observations (rows) and variables (columns) are in the `mpg` dataset?

```{r}



```

4. Select the `manufacturer`, `model` and `year` columns from the `mpg` dataset.

```{r}



```

5. Identify the subset of cars/rows where city fuel economy (`cty`) is greater than 20 and highway fuel economy (`hwy`) is greater than 30. Assign this output to an object called `eff`. How many cars/rows are present?

```{r}



```

6. How many fuel efficient cars (in the `eff` dataset) were manufactured in the year 1999?

```{r}



```

7. Filter cars from the overall `mpg` dataset that do not have 4 cylinder engines. How many cars/rows are there?

```{r}



```

8. Filter cars to only those in the "suv" or "minivan" class. How many cars/rows are there?

```{r}



```

9. Filter cars with displacements (`displ`) greater than 4 and that are all 4 wheel drive (`drv`) (a value of 4 for `drv`). How many cars/rows are there. 

```{r}

```

