---
title: "Lab: Summarizing Data"
output: html_document
editor_options: 
  chunk_output_type: console
# Adapted from Carrie Wright and Ava Hoffman
---

This lab uses public transport data from Baltimore, used in Lab2. The .csv file is located in `data/Charm_City_Circulator_Ridership.csv`

```{r read_in_data}
library(tidyverse)


```

1.  How many days worth of data are in the data set? You can assume each observation/row is a different day

```{r q1}

```

2.  What is the total (sum) number of boardings on the green bus (`greenBoardings`)?

```{r q2}

```

3.  How many days are missing daily ridership? Use `is.na()` and `sum()`.

```{r q3}

```

4.  Group the data by day of the week (`day`) and calculate the sum for riders.

```{r q4}

```

## **Extra practice:**

5.  Describe the distribution of `orangeBoardings`(use `quantile()`). The `quantile()` function works like `mean()`, `sum()`, etc.

```{r q6}

```

6.  Describe the distribution of `orangeBoardings`(use `quantile()`), but this time stratify by day of the week.

```{r q7}

```
