---
title: "HW 1: Cleaning UFO Data"
output: html_document
# Elements adapted from Carrie Wright and Ava Hoffman
---

In this lab you will have more fun with the (messy) UFO sighting data set.

# Cleaning UFO Sightings!

```{r}
library(tidyverse)
```

1.  Read in the UFO dataset as an R object called `ufo`. You can read it from: `data/ufo/ufo_data_complete.csv` . You can ignore the "problems" with some rows.

```{r}
# You can also grab it from here: https://raw.githubusercontent.com/SISBID/Module1/gh-pages/data/ufo/ufo_data_complete.csv

```

### Stringr Exercises

2.  Clean up the column/variable names of the `ufo` dataset to remove spaces and non-alphanumeric characters. You can use the `dplyr::rename()` function or look into the `janitor::clean_names()` function. (Remember you may need to install janitor in addition to tidyverse.)

```{r}

```

3.  How many UFO sightings were reported on a time scale of minutes, specifically using the `duration (hours/min)` originally-named column? (hint: explore the various entries in this column and then use a `stringr` function to match a flexible pattern. you can ignore observations like "1/2 hour" and similar ones that don't contain some version of the word "minutes")

```{r}

```

4.  How accurate is the (formerly-named) `duration (seconds)` column? I.e. how many of the above minutes-scale observations have durations less than 120 seconds or greater than 3800 seconds?

```{r}

```

5.  How many ufo sighting cities end in (a) "field", (b) "ton" and (c) "port"? Hint: See regular expressions to help feed your stringr functions (aka regex)

```{r}

```

6.  How many ufo sighting comments reference the "greys" ?

```{r}

```

7. Which cities did these occur in?

```{r}

```

### Practice with lubridate

8. How many encounters occurred in the year 2010? This question might be tricky to answer until we convert our table columns into something more workable. What would make the `datetime` column format more useful? 
Create a new column that converts the `datetime` column into a lubridate `<date>` format. Then answer the question.  Call this new column `date_enc`.


```{r}

```


9. Convert the other "date/time" column `date_posted` to a lubridate formate also. Call this column `date_posted2`
Hint/note: may need to use "quiet = TRUE" for some of your lubridate conversions to avoid errors.

```{r}

```

9b. Create new separate columns for `year`, `month`, `day`, and `hour`, based on the `date_enc`.
Hint/note: may need to use "quiet = TRUE" for some of your lubridate conversions to avoid errors.

```{r}

```

(extra credit) 10. Let's take a look at the time it took for these alleged encounters to be reported. Compare the encounter date/time to the posted date.  What is the mean durations between the two?
Hint: maybe take a look here: https://lubridate.tidyverse.org/reference/time_length.html 

```{r}

```

11.  Can you generate a count summary over the years for the encounters? And over the years posted? (using the counts() function)

```{r}

```

(extra credit) 12. Can you plot in a barplot of the encounter counts ?
Hint: You might need to remove NAs
```{r}

```

### Plotting Review Fun

12.  Can you make a scatter plot of all encounters after 2013 occurring in the us?

```{r}
# Hint: may want to use to reduce clutter/speed up rendering.
# + theme(axis.text.x=element_blank(),  #remove x axis labels                
#         axis.ticks.x=element_blank(), #remove x axis ticks          
#         axis.text.y=element_blank(),  #remove y axis labels           
#         axis.ticks.y=element_blank()) #remove y axis ticks     )`


```

### Practice with Forcats

<https://forcats.tidyverse.org/>

11. What are the most common `shapes` of ufos for the `state` of Texas?  (first try with counts())
```{r}

```

12. Can you make barplot displaying this information? (this time using forcats)

```{r}

```

13. Can you make the barplot ordered by counts?

```{r}

```

14. This might display too many items for a concise plot. Can you reduce the count categories and replot?

```{r}

```


15. What was the most common shape for Texas?  How about Texas in the year 2010?

```{r}

```

